---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# malaviR

The goal of this package is to be an R interface to the global avian haemosporidian database MalAvi (http://mbio-serv2.mbioekol.lu.se/Malavi/). The package includes functions for downloading data from the MalAvi website directly into your R environment. You can also use malaviR to BLAST your haemosporidian DNA sequences against the MalAvi database programmatically, which should facilitate comparisons with MalAvi.  
  
The package also includes a key for linking the host taxonomic classifications in MalAvi with the avian taxonomic classifications found on http://birdtree.org/.

## Installation

You can install malaviR from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("vincenzoaellis/malaviR")
```

## Download tables from MalAvi

There are nine tables that can be downloaded from the MalAvi website summarizing the dataset. Here's how you would download the table of references from the MalAvi database:

```{r example}
library(malaviR)
refs <- extract_table("Table of References")
head(refs) # check it out
```

The `extract_table()` help file lists all nine tables you can download directly into `R`. Or you can specify `"all"` and get them all as a list.  

## BLAST a sequence to MalAvi

You can BLAST your own sequences against the MalAvi database using the `blast_malavi()` function. Your input sequence just needs to be specified as a character string.  

```{r example2}
## define a sequence. This is the Plasmodium parasite ACAGR1
ACAGR1 <- "GCAACTGGTGCTTCATTTGTATTTATTTTAACTTATTTACATATTTTAAGAGGATTAAATTATTCATATTCATATTTACCTTTATCATGGATATCTGGATTAATAATATTTTTAATATCTATAGTAACAGCTTTTATGGGTTACGTATTACCTTGGGGTCAAATGAGTTTCTGGGGTGCTACCGTAATAACTAATTTATTATATTTTATACCTGGACTAGTTTCATGGATATGTGGTGGATATCTTGTAAGTGACCCAACCTTAAAAAGATTCTTTGTACTACATTTTACATTTCCTTTTATAGCTTTATGTATTGTATTTATACATATATTCTTTCTACATTTACAAGGTAGCACAAATCCTTTAGGGTATGATACAGCTTTAAAAATACCCTTCTATCCAAATCTTTTAAGTCTTGATATTAAAGGATTTAATAATGTATTAGTATTATTTTTAGCACAAAGTTTATTTGGAATACT"

## BLAST it against the MalAvi database and save the top five hits to a data frame
hits <- blast_malavi(ACAGR1)
hits # check it out
```

## Download the MalAvi sequence alignments

Using the `extract_alignment()` function, you can download all of the sequences on MalAvi, or the complete or nearly complete sequences only, or the sequences associated with morphological species only. These alignments will appear as objects of the class `DNAbin`, defined in the `ape` package.

```{r example3}
## download all sequences
all.seqs <- extract_alignment("all seqs")
all.seqs #check it out
```

The `extract_alignment()` help file lists the names of the sequence alignments that the function understands.
