---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# malaviR

The goal of this package is to be an R interface to the global avian haemosporidian database MalAvi (http://mbio-serv2.mbioekol.lu.se/Malavi/). The package includes functions for downloading data from the MalAvi website directly into your R environment. You can also use malaviR to BLAST your haemosporidian DNA sequences against the MalAvi database programmatically, which should facilitate comparisons with MalAvi. 
The package also includes a key for linking the host taxonomic classifications in MalAvi with the avian taxonomic classifications found on http://birdtree.org/.

## Installation

You can install malaviR from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("vincenzoaellis/malaviR")
```

## Download tables from MalAvi

Here's how you would download the table of references from the MalAvi database:

```{r example}
library(malaviR)
refs <- extract_table("Table of References")
head(refs) # check it out
```

## BLAST a sequence to MalAvi

```{r example2}
## define a sequence. This is the Plasmodium parasite ACAGR1
ACAGR1 <- "GCAACTGGTGCTTCATTTGTATTTATTTTAACTTATTTACATATTTTAAGAGGATTAAATTATTCATATTCATATTTACCTTTATCATGGATATCTGGATTAATAATATTTTTAATATCTATAGTAACAGCTTTTATGGGTTACGTATTACCTTGGGGTCAAATGAGTTTCTGGGGTGCTACCGTAATAACTAATTTATTATATTTTATACCTGGACTAGTTTCATGGATATGTGGTGGATATCTTGTAAGTGACCCAACCTTAAAAAGATTCTTTGTACTACATTTTACATTTCCTTTTATAGCTTTATGTATTGTATTTATACATATATTCTTTCTACATTTACAAGGTAGCACAAATCCTTTAGGGTATGATACAGCTTTAAAAATACCCTTCTATCCAAATCTTTTAAGTCTTGATATTAAAGGATTTAATAATGTATTAGTATTATTTTTAGCACAAAGTTTATTTGGAATACT"

## BLAST it against the MalAvi database and save the top five hits to a data frame
hits <- blast_malavi(ACAGR1)
hits # check it out
```
