---
title: "Using malaviR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using malaviR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(malaviR)
```

`malaviR` gives you access to the [MalAvi](http://mbio-serv2.mbioekol.lu.se/Malavi/)
database of avian haemosporidian parasite lineages. Because MalAvi is no longer a
live website, the database is **bundled with the package** as a dated snapshot and
everything below works offline.

## Which database version am I using?

```{r}
malavi_version()
malavi_versions()
```

## The data tables

There are five tables; ask for one by name (or `"all"` for a list of all five):

```{r}
hosts <- extract_table("Hosts and Sites Table")
dim(hosts)
colnames(hosts)[1:6]
```

## The sequence alignment

`extract_alignment()` returns the cytochrome *b* alignment as an `ape::DNAbin`
object, optionally subset by parasite genus:

```{r}
aln <- extract_alignment()
aln
plasmodium <- extract_alignment(genus = "Plasmodium")
plasmodium
```

## Repeated haplotypes (synonymies)

Different MalAvi names are sometimes given to the same sequence. This inflates
estimates of parasite diversity, an issue documented by
[Tamayo-Quintero et al. (2025)](https://doi.org/10.1371/journal.ppat.1012911).
`synonymy_report()` quantifies it and lists the lineages to investigate:

```{r}
report <- synonymy_report(method = "strict")
report$summary
head(report$synonymies)
```

The default `method = "overlap"` is more thorough — it also collapses short,
partial sequences that are identical to a longer one over the region they
overlap — but is slower on the full alignment.

`clean_alignment()` produces a de-duplicated alignment. It keeps the most complete
sequence in each group by default, and you can override that with `keep`:

```{r}
clean <- clean_alignment(aln, method = "strict")
length(clean$dropped)
clean$alignment_clean
```

## Local BLAST-like search

`blast_malavi()` searches a query sequence against the bundled alignment and
returns the most similar lineages. It uses the Bioconductor packages **DECIPHER**
(>= 3.0) and **Biostrings**, so it is shown here but not run:

```{r, eval = FALSE}
query <- gsub("-", "", paste(as.character(aln[1, ]), collapse = ""))
blast_malavi(query, top_n = 5)
```

## Host taxonomy

`match_taxonomy()` aligns host species names to the modern
[clootl](https://github.com/McTavishLab/clootl) (eBird) avian taxonomy. Names that
cannot match a single species (synonyms, hybrids, `sp.`) are flagged rather than
forced:

```{r}
res <- match_taxonomy(c("Turdus merula", "Cyanistes caeruleus", "Anas sp."))
res$key
```

The pre-built crosswalk for every MalAvi host is bundled as `taxonomy`:

```{r}
data(taxonomy)
head(taxonomy)
table(taxonomy$match_type)
```

## Working with phylogenies

`clean_names()` strips the genus prefix from alignment tip labels so they match
the data tables, and `sister_taxa()` returns the sister tips at a node:

```{r}
clean_names(c("H_COLL2", "P_GRW4", "L_CIAE02"))

tree <- ape::read.tree(text = "((A,B),(C,(D,E)));")
sister_taxa(tree, node = 8)
```
